---
title: "Aligning Sample To Peptides"
author: "Simon Hickinbotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction


This tutorial will show you how to load three replicates of a sample from files and align them with some peptides. We'll show the basic process first, and then give some details about how you would go about 

# How to load some data

The bacollite package comes with some data files that we can use during this example, but we need some R code to find where they are on our system. First we need R to tell us the folder that the data is in: 

```{r}
fpath <- system.file("extdata",package="bacollite")
```

The files that we are looking for are:

```
20150326_SF_YG65_94_I11.txt  
20150326_SF_YG65_94_I14.txt  
20150326_SF_YG65_94_I17.txt
```
You can see that the last part of the file name before the "." is the spot reference, and there are three spots: `I11`, `I14` and `I17`. We need this information to use the `load.sample` function to load the three samples into R, like this: 

```{r}
library(bacollite)
froot = sprintf("%s/20150326_SF_YG65_94_",fpath)
samples <- load.sample(froot = froot,spots = c("I11","I14","I17"),fext=".txt")
```









# Load some peptides







# Run the alignment








# Perform a classification

























# Sample data formats

The example data files are tab-delimited text files with two coulmns. A small portion of one of them is shown here:

```
2019.60878627237	21
2019.63791216278	16
2019.66703826364	3
2019.69616457538	2
2019.72529109758	6
2019.75441783023	5
2019.78354477354	17
```

The first column is the *mass* and the second column is the *intensity* (i.e. the ion count) for the MALDI sample. 



# Peptide files

See the vignette


# Processing entire datasets




-----

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
